---
title: "Introduction to SentimentAnalysis"
author: 
- "Stefan Feuerriegel"
- "Nicolas Proellochs"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to SentimentAnalysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

**SentimentAnalysis** performs a sentiment analysis of textual contents in R. This implementation utilizes various existing dictionaries, such as General Inquirer, Harvard IV or Loughran-McDonald. Furthermore, it can also create customized dictionaries. The latter uses LASSO regularization as a statistical approach to select relevant terms  based on an exogeneous response variable. 

## Install
Using the **devtools** package, you can easily install the latest development version of **SentimentAnalysis** with

```{r,eval=FALSE}
library(devtools)

# Option 1: download and install latest version from GitHub
install_github("sfeuerriegel/SentimentAnalysis")

# Option 2: install directly from bundled archive
install_local("SentimentAnalysis_1.0.0.tar.gz")
```

Note: that you have to specify the path either to the directory of **SentimentAnalysis** or to the bundled archive **SentimentAnalysis_1.0.0.tar.gz**

## Usage

This section shows the basic functionality to crawl for ad hoc filings. The following lines extract the ad hoc disclosure that was published the latest. 

```{r}
library(SentimentAnalysis)
```

## Brief demonstration

```{r}
# Analyze a single string to obtain a binary response (positive / negative)
sentiment <- analyzeSentiment("Yeah, this was a great soccer game of the German team!")
convertToBinaryResponse(sentiment)$SentimentGI
```

```{r}
# Create a vector of strings
documents <- c("Wow, I really like the new light sabers!",
               "That book was excellent.",
               "R is a fantastic language.",
               "The service in this restaurant was miserable.",
               "This is neither positive or negative.",
               "The waiter forget about my a dessert -- what a poor service!")

# Analyze sentiment
sentiment <- analyzeSentiment(documents)

# Extract dictionary-based sentiment according to the Harvard-IV dictionary
sentiment$SentimentGI

# View sentiment direction (i.e. positive, neutral and negative)
convertToDirection(sentiment$SentimentGI)

response <- c(+1, +1, +1, -1, 0, -1)

#TODO: compareToResponse(sentiment, response)

#TODO: plotSentimentResponse(sentiment$SentimentGI, response)
```

The **SentimentAnalysis** package works very clever and neatly here in order to remove the effort
for the user: it recognizes that the user has inserted
a vector of strings and thus automatically performs a set of default preprocessing operations from
text mining. Hence, it tokenizes each document and finally converts the input into a 
document-term matrix. All of the previous operations are undertaken without manual specification. 
The **analyzeSentiment** routine also accepts other input formats in case the user has already
performed a preprocessing or wants to specify a specific set of operations.

## Interface

dtm
VectorSource
stirngs

## Built-in dictionaries

The **SentimentAnalysis** package entails three different dictionaries:
* Harvard-IV dictionary as used in the General Inquirer program
* Henry's Financial dictionary
* Loughran-McDonald Financial dictionary
All of them can be manually inspected and even accessed as follows:

```{r}
# Make dictionary available in the current R environment
data(DictionaryGI)
# Display the internal structure 
str(DictionaryGI)
# Access dictionary as an object of type SentimentDictionary
dict.GI <- loadDictionaryGI()
# Print summary statistics of dictionary
summary(dict.GI)

data(DictionaryHE)
str(DictionaryHE)

data(DictionaryLM)
str(DictionaryLM)
```

## Sentiment analysis with built-in dictionaries

## Dictionary generation

```{r}
# Create a vector of strings
documents <- c("This is a good thing!",
               "This is a very good thing!",
               "This is okay.",
               "This is a bad thing.",
               "This is a very bad thing.")
response <- c(1, 0.5, 0, -0.5, -1)

# Generate dictionary with LASSO regularization
dict <- generateDictionary(documents, response)

dict

summary(dict)

#TODO
# compareDictionaries(dict,
#                     loadDictionaryGI())
# 
# sentiment <- predict(dict, documents)
# compareToResponse(sentiment, response)
# plotSentimentResponse(sentiment, response)
```

```{r}
test_documents <- c("This is neither good nor bad",
                    "What a good idea!",
                    "Not bad")
test_response <- c(0, 1, 1)

pred <- predict(dict, test_documents)

compareToResponse(pred, test_response)
plotSentimentResponse(pred, test_response)

# TODO: compare to static dictionary
```

## Performance evaluation

## Visualizations

## Aggregation of documents

## Worked examples

Reuters oil

```{r}
library(tm)
data(crude)

# Analyze sentiment
sentiment <- analyzeSentiment(crude)

# Count positive and negative news releases
table(convertToBinaryResponse(sentiment$SentimentLM))

# News releases with highest and lowest sentiment
crude[[which.max(sentiment$SentimentLM)]]$meta$heading
crude[[which.min(sentiment$SentimentLM)]]$meta$heading

# View summary statistics of sentiment variable
summary(sentiment$SentimentLM)

# Visualize distribution of standardized sentiment variable
hist(scale(sentiment$SentimentLM))

# Compute cross-correlation 
cor(sentiment[, c("SentimentLM", "SentimentHE", "SentimentGI")])

# TODO: crude oil returns 1987-02-26 until 1987-03-02 with aggregate
datetime <- do.call(c, lapply(crude, function(x) x$meta$datetimestamp))
plotSentiment(sentiment$SentimentLM)
plotSentiment(sentiment$SentimentLM, x=datetime, cumsum=TRUE)
```

ngram

standardize and intercept

aggregate

library(tm)
## References

For further information refer to

* [TODO DictGen](http://www.dgap.de/dgap/News/?newsType=ADHOC)

## License

**SentimentAnalysis** is released under the [MIT License](https://opensource.org/licenses/MIT)
Copyright (c) 2016 Stefan Feuerriegel & Nicolas Pröllochs
